The Multi-Stencil Language

### 2016-05-20 ARTICLES/IJPP/article/article.tex
IJPP article modifications parties 4 5 et 6

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Contributions :
\begin{itemize}
\item Formalism of a multi-stencil program and its parallelization
\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
\item Evaluation of the different types of parallelism introduced + compiler + fusion
\end{itemize}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
%------------------------------------------------------------------------------
\section{Static Scheduling}
\label{sect:msp}
\input{msp.tex}
% From the dependency graph we study a first dump: static scheduling
% \begin{itemize}
% \item description of the algorithm to get the binary tree decomposition of a minimal serie-parallel graph
% \item use of the serie-parallel tree decomposition to detect fusions of kernels
% \item example (the same one) from the dependency graph to the binary tree
% \end{itemize}
%------------------------------------------------------------------------------
\section{MSL Compiler}
\label{sect:comp}
\input{compiler.tex}
% \begin{itemize}
% \item Overall algrotihm~\ref{alg:compiler}
% \item Big example from the language to the static sheduling/optimization~\ref{fig:example}
% \end{itemize}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\begin{itemize}
\item short description of the implementation (compiler in Python, DDS = SkelGIS, Components)
\item evaluation of the compiler (execution times of the different steps)
\item evaluation of the data parallelism (weak/strong scaling)

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
  \caption{weak-scaling 400x400 blocks.}
  \label{fig:mesh}
\end{center}\end{figure}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
  \caption{weak-scaling 800x800 blocks}
  \label{fig:mesh}
\end{center}\end{figure}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
  \caption{strong-scaling 10kx10k, 1k iterations}
  \label{fig:mesh}
\end{center}\end{figure}

\item evaluation of the fusion

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
  \caption{strong scaling 500x500, 200 iterations, with and without fusion}
  \label{fig:mesh}
\end{center}\end{figure}

\item evaluation of the hybrid parallelism

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
  \caption{MPI only vs MPI+threads with \emph{close} scheduling policy (OpenMP)}
  \label{fig:mesh}
\end{center}\end{figure}

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
  \caption{MPI only vs MPI+threads with \emph{spread} scheduling policy (OpenMP)}
  \label{fig:mesh}
\end{center}\end{figure}

\item analytic model for the results

\begin{figure}[!h]\begin{center}
  \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
  \caption{Computation vs communication times in the MPI only application}
  \label{fig:mesh}
\end{center}\end{figure}

\end{itemize}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\begin{itemize}
\item Cartesian meshes : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
\item Unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
\item Halide for pipeline of stencil computations in image (Cartesian) = mesh dependent
\item SYCL same Halide mais mesh dependent
\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
\end{itemize}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
%------------------------------------------------------------------------------

>>>>

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Contributions :
\begin{itemize}
\item Formalism of a multi-stencil program and its parallelization
\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
\item Evaluation of the different types of parallelism introduced + compiler + fusion
\end{itemize}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
%------------------------------------------------------------------------------
\section{Static Scheduling and performance model}
\label{sect:msp}
\input{msp.tex}
%------------------------------------------------------------------------------
\section{Implementation}
\label{sect:comp}
\input{compiler.tex}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\begin{itemize}
\item evaluation of the compiler (execution times of the different steps)
\item evaluation of the data parallelism (weak/strong scaling)
\item evaluation of fusions
\item evaluation of the hybrid performance
\item validation of the performance model
\end{itemize}

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/400/median_weak.pdf}}
%   \caption{weak-scaling 400x400 blocks.}
%   \label{fig:mesh}
% \end{center}\end{figure}

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{../results/weak_scaling/800/median_weak.pdf}}
%   \caption{weak-scaling 800x800 blocks}
%   \label{fig:mesh}
% \end{center}\end{figure}

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{../results/strong_scaling/10K_1K/median_strong.pdf}}
%   \caption{strong-scaling 10kx10k, 1k iterations}
%   \label{fig:mesh}
% \end{center}\end{figure}

% \item evaluation of the fusion

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/fusVSbase.pdf}}
%   \caption{strong scaling 500x500, 200 iterations, with and without fusion}
%   \label{fig:mesh}
% \end{center}\end{figure}

% \item evaluation of the hybrid parallelism

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_close_median.pdf}}
%   \caption{MPI only vs MPI+threads with \emph{close} scheduling policy (OpenMP)}
%   \label{fig:mesh}
% \end{center}\end{figure}

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/base_spread_median.pdf}}
%   \caption{MPI only vs MPI+threads with \emph{spread} scheduling policy (OpenMP)}
%   \label{fig:mesh}
% \end{center}\end{figure}

% \item analytic model for the results

% \begin{figure}[!h]\begin{center}
%   \resizebox{8cm}{!}{\includegraphics{../results/task_scaling/500_200/analytic/times.pdf}}
%   \caption{Computation vs communication times in the MPI only application}
%   \label{fig:mesh}
% \end{center}\end{figure}

% \end{itemize}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\begin{itemize}
\item stencil compilers cartesian : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
\item stencil compilers unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
\item Halide and Nabla for pipeline of stencil computations in image (Cartesian) = mesh dependent
\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
\end{itemize}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
%------------------------------------------------------------------------------

### 2016-05-20 ARTICLES/IJPP/article/compiler.tex
IJPP article modifications parties 4 5 et 6

The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}. First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computations. Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$. Third, N-Shapes are removed from $\Gamma_{dep}$ to get a MSPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$ (line 6). Lines 7 to 9 are explained in the next section. Finally a dump of $TSP$ is performed by the compiler and will be described in Section~\ref{sect:eval}.

\begin{algorithm}
\caption{MSL Compiler}
\label{alg:compiler}
\begin{algorithmic}[1]
\Procedure{compileMSL} {file}
\\$\Gamma$ = parser(file)
\\$\Gamma_{sync} = F_{sync}(\Gamma,0)$
\\$\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
\\removeNSH($\Gamma_{dep}$)
\\$TSP = BDT(L^{-1}(\Gamma_{dep}))$
\\$TSP = F_{tsp}(TSP,root(TSP))$
\If {data parallel}
\\$TSP = F_{tsp}(TSP)$
\EndIf
\\dump(TSP)
\EndProcedure
\end{algorithmic}
\end{algorithm}

%--------------------
\subsection{Canonical form and fusions}

The serie-parallel binary tree decomposition $TSP$ obtained from $BDT(L^{-1}(\Gamma_{dep}))$ can be reduced to its canonical form which consists in recursively merging successive $S$ vertices or successive $P$ vertices.

\medskip
\noindent \textbf{Definitions} For a tree $T(V,E)$
%\begin{mydef}
\begin{itemize}
\item we denote $root(T)$, the function which returns the root vertex of the tree $T$;
\item we denote $type(v)$ with $v \in V$ the function which returns the type of a vertex $v$. This type is either $S$, $P$, $comp$ or $sync$;
\item We denote $replace(E,v_1,v_2)$, the function which replaces for all edges of the set $E$, $v_1$ by $v_2$.
\end{itemize}

\begin{mydef}
We denote as a sequence or parallel sub-tree $T(V,E)$ sub-trees for which the root vertex respectively verifies $type(root(T))=S$ or $type(root(T))=P$.
\end{mydef}

\begin{myth}
\begin{itemize}
\item A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
\begin{equation*}
T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
\end{equation*}
\item A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
\begin{equation*}
T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
\end{equation*}
\end{itemize}
\end{myth}

These reductions applied recursively on the serie-parallel binary tree decomposition of $\Gamma_{dep}$ produce the canonical serie-parallel tree decomposition 
\begin{equation*}
TSP = F_{tsp}(TSP,root(TSP),successors(root(TSP)).
\end{equation*}
The recursive function $F_{tsp}$ is based on a Depth First Search algorithm and defined by Algorithm~\ref{alg:tsp}, where $E_{rooted(s)}$ represents the set of edges of the sub-tree rooted by the vertex $s$.

\begin{algorithm}
\caption{$F_{tsp}$}
\label{alg:tsp}
\begin{algorithmic}[1]
\Procedure{$F_{tsp}$} {$TSP$,root,succ}
\For {$s \in succ$}
\If {type($s$)=$S$ or type($s$)=$P$}
\\$F_{tsp}(TSP,s,successors(s))$
\If {type($s$)=type(root)}
\\$TSP = (V_{tsp} \setminus s,E_{tsp} \cup replace(E_{rooted(s)},s,root))$
\EndIf
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

Using MSL, it is possible to ask for data parallelization of the application, or for an hybrid parallelization. Even though the MSL language is not dedicated to produce very optimized stencil codes, but to produces the parallel pattern of the application, building the $TSP$ tree make available an easy optimization when the data parallelization technique is the only one used. This optimization consists in proposing a valid merge of some computation kernels inside a single space loop. As a result, the user can use this valid fusion of kernels or not when implementing those. The fusion function $F_{fus}$ is described in Algorithm~\ref{alg:fus}, where $parent(k)$ returns the parent vertex of $k$ in the tree, and where $k_{i;j}$ represents the fusion of $k_i$ and $k_j$ keeping the sequential order $i;j$ if $i$ is computed before $j$ in $TSP$.

\begin{algorithm}
\caption{$F_{fus}$}
\label{alg:fus}
\begin{algorithmic}[1]
\Procedure{$F_{fus}$} {$TSP(V,E)$}
\For {$(k_i,k_j) \in V^2$}
\If {$type(k_i)=type(k_j)=comp$}
\If {parent($k_i$)==parent($k_j$)}
\If {parent($k_i$)==$S$}
\If {$D_i==D_j$}
\\propose the fusion $k_{i;j}$
\EndIf
\EndIf
\If {parent($k_i$)==$P$}
\If {$D_i==D_j$ and $R_i \cap R_j \neq \emptyset$}
\\propose the fusion $k_{i;j}$
\EndIf
\EndIf
\EndIf
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

We are not arguing that such a simple fusion algorithm could be as good as complex cache optimization techniques which can be found in stencil DSLs for example~\cite{spaaTangCKLL11}. However, this fusion takes place at a different level and can bring performance improvments as it will be illustrated in Section~\ref{sect:eval}. This fusion algorithm relies on very simple statements:
\begin{itemize}
\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $S$ in TSP are, by construction, data dependant. As a result, what is written by the first one is read by the second one. Thus, at least one data is common to those computations (the one written by $k_i$). Thus, if the computation domains verify $D_i=D_j$, the fusion of $k_i$ and $k_j$ will decrease cache misses.
\item Two successive computation kernels $k_i$ and $k_j$ which are under the same parent vertex $P$ in TSP are not, by construction, data dependant. However, if the computation domains verify $D_i=D_j$, and if $R_i \cap R_j \neq \emptyset$ cache misses could also be decreased by the fusion $k_{i;j}$.
\end{itemize}

\todo{solve order problem in $\Gamma_{dep}$ and $TSP$}

%--------------------
\subsection{Overall example}

% Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time. 
A complete example of reduction is given in Figure~\ref{fig:example}. The initial $\Gamma_{hybrid}$ DAG and its inverse line graph are given first, followed by the set of reductions. The label of the final single edge corresponds to the binary tree decomposition of the initial $\Gamma_{hybrid}$ graph.

%====BIG EXAMPLE
\begin{figure}[ht!]
\begin{center}
% INITIAL GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$c_0$};
   \node[] (star1) at (1,0) {$*_1$};
   \node[] (c1) at (2,0) {$c_1^*$};
   \node[] (c2) at (3,0.5) {$c_2$};
   \node[] (star4) at (3,1.5) {$*_4$};
   \node[] (c3) at (3,-0.5) {$c_3$};
   \node[] (c4) at (4,0.5) {$c_4^*$};
   \node[] (c5) at (4,-0.5) {$c_5$};
   \node[] (c6) at (5,0.5) {$c_6$};
   \node[] (c7) at (6,0) {$c_7$};
   \node[] (star8) at (7,0) {$*_8$};
   \node[] (c8) at (8,0) {$c_8^*$};

  \path[->]
    (c0) edge node {} (star1)
    (star1) edge node {} (c1)
    (c1) edge node {} (c2)
         edge node {} (c3)
         edge node {} (star4)
    (star4) edge node {} (c4)
    (c2) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (star8)
    (star8) edge node {} (c8);
  \end{tikzpicture}
}
\\
% INVERSE LINE GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c0d) at (1,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (star1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (7,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (star8d) at (8,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

  \path[->]
    (c0s) edge node [above] {$c_0$} (c0d)
    (c0d) edge node [above] {$*_1$} (star1d)
    (star1d) edge node [above] {$c_1^*$} (c1d)
    (c1d) edge node [above] {$c_2$} (c2d)
          edge node [above] {$c_3$} (c3d)
          edge [bend left=80] node [above] {$*_4$} (c2d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {$c_7$} (c7d)
    (c7d) edge node [above] {$c_8$} (star8d)
    (star8d) edge node [above] {$c_8$} (c8d)
    (c3d) edge node [above] {$c_5$} (c6d);
  \end{tikzpicture}
}
\\
% TRANSFORMATIONS
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {$c_2$} (c2d)
          edge node [above] {$c_3$} (c3d)
          edge [bend left=80] node [above] {$*_4$} (c2d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c3d) edge node [above] {$c_5$} (c6d)

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

   %reduction c3 c5
   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (4,-1) {$c_3$};
   \node[] (c5) at (5,-1) {$c_5$};

   %reduction c2 *4
   \node[] (p0) at (3.5,1) {$\mathcal{P}$};
   \node[] (c2) at (3,2) {$c_2$};
   \node[] (star4) at (4,2) {$*_4$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {} (c2d)
    (c2d) edge node [above] {$c_4^*$} (c4d)
    (c4d) edge node [above] {$c_6$} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c1d) edge [bend right] node [above] {} (c6d)

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8)

    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)

    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4);
  \end{tikzpicture}
}
\end{center}
\caption{Big example}
\label{fig:example}
\end{figure}

\begin{figure}[ht!]
\ContinuedFloat
\begin{center}
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

   %reduction c3 c5
   \node[] (s5) at (4.5,-2) {$\mathcal{S}$};
   \node[] (c3) at (4,-1) {$c_3$};
   \node[] (c5) at (5,-1) {$c_5$};

   %reduction c2 c6
   \node[] (s6) at (4.5,1) {$\mathcal{S}$};
   \node[] (p0) at (3.5,2) {$\mathcal{P}$};
   \node[] (c2) at (3,3) {$c_2$};
   \node[] (star4) at (4,3) {$*_4$};
   \node[] (s7) at (5.5,2) {$\mathcal{S}$};
   \node[] (c4) at (5,3) {$c_4^*$};
   \node[] (c6) at (6,3) {$c_6$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge [bend left] node [above] {} (c6d)
    (c6d) edge node [above] {} (c8d)
    (c1d) edge [bend right] node [above] {} (c6d)

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8)

    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)

    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (7,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   %reduction c0 c1
   \node[] (s0) at (1.5,0.5) {$\mathcal{S}$};
   \node[] (s1) at (1,1.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,2.5) {$c_0$};
   \node[] (star1) at (1.5,2.5) {$*_1$};
   \node[] (c1) at (2,1.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,0.5) {$\mathcal{S}$};
   \node[] (s3) at (7,1.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,2.5) {$c_7$};
   \node[] (star8) at (7.5,2.5) {$*_8$};
   \node[] (c8) at (8,1.5) {$c_8$};

   \node[] (p1) at (4.5,0.5) {$\mathcal{P}$};
   %reduction c3 c5
   \node[] (s5) at (5.5,1.5) {$\mathcal{S}$};
   \node[] (c3) at (5,2.5) {$c_3$};
   \node[] (c5) at (6,2.5) {$c_5$};
   %reduction c2 c6
   \node[] (s6) at (3.5,1.5) {$\mathcal{S}$};
   \node[] (p0) at (2.8,2.5) {$\mathcal{P}$};
   \node[] (c2) at (2.3,3.5) {$c_2$};
   \node[] (star4) at (3.3,3.5) {$*_4$};
   \node[] (s7) at (4.2,2.5) {$\mathcal{S}$};
   \node[] (c4) at (3.8,3.5) {$c_4^*$};
   \node[] (c6) at (4.8,3.5) {$c_6$};

  \path[->]
    (c0s) edge node {} (c1d)
    (c1d) edge node [above] {} (c6d)
    (c6d) edge node [above] {} (c8d)

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8)

    (p1) edge node {} (s6)
         edge node {} (s5)
    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)
    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

   \node[] (s8) at (4.5,0.5) {$\mathcal{S}$};
   \node[] (s9) at (3,1.5) {$\mathcal{S}$};

   %reduction c0 c1
   \node[] (s0) at (1.5,2.5) {$\mathcal{S}$};
   \node[] (s1) at (1,3.5) {$\mathcal{S}$};
   \node[] (c0) at (0.5,4.5) {$c_0$};
   \node[] (star1) at (1.5,4.5) {$*_1$};
   \node[] (c1) at (2,3.5) {$c_1^*$};

   %reduction c7 c8
   \node[] (s2) at (7.5,1.5) {$\mathcal{S}$};
   \node[] (s3) at (7,2.5) {$\mathcal{S}$};
   \node[] (c7) at (6.5,3.5) {$c_7$};
   \node[] (star8) at (7.5,3.5) {$*_8$};
   \node[] (c8) at (8,2.5) {$c_8$};

   \node[] (p1) at (4.5,2.5) {$\mathcal{P}$};
   %reduction c3 c5
   \node[] (s5) at (5.5,3.5) {$\mathcal{S}$};
   \node[] (c3) at (5,4.5) {$c_3$};
   \node[] (c5) at (6,4.5) {$c_5$};
   %reduction c2 c6
   \node[] (s6) at (3.5,3.5) {$\mathcal{S}$};
   \node[] (p0) at (2.8,4.5) {$\mathcal{P}$};
   \node[] (c2) at (2.3,5.5) {$c_2$};
   \node[] (star4) at (3.3,5.5) {$*_4$};
   \node[] (s7) at (4.2,4.5) {$\mathcal{S}$};
   \node[] (c4) at (3.8,5.5) {$c_4^*$};
   \node[] (c6) at (4.8,5.5) {$c_6$};

  \path[->]
    (c0s) edge node {} (c8d)

    (s8) edge node {} (s9)
         edge node {} (s2)
    (s9) edge node {} (p1)
         edge node {} (s0)

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8)

    (p1) edge node {} (s6)
         edge node {} (s5)
    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)
    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\end{center}
\caption{Big example}
\label{fig:example}
\end{figure}

% At this point, the $\mathcal{MSP}$ program can be dumped to any parallel language in which a sequence and a parallel execution are available. For example, we could imagine a parallel functionnal language with a binary \emph{sequence} function and a binary \emph{parallel} function.

>>>>

The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}. First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computation kernels. Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$. Third, N-Shapes are removed from $\Gamma_{dep}$ to get a MSPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$ (line 6). If only a data parallelization has been chosen, Line 8 performs the fusion of kernels. Finally a dump of $TSP$ is performed by the compiler. Implementation choices of the dump are described in the next subsection.

\begin{algorithm}
\caption{MSL Compiler}
\label{alg:compiler}
\begin{algorithmic}[1]
\Procedure{compileMSL} {file}
\State $\Gamma$ = parser(file)
\State $\Gamma_{sync} = F_{sync}(\Gamma,0)$
\State $\Gamma_{dep} = F_{dep}(\Gamma_{sync},0)$
\State removeNSH($\Gamma_{dep}$)
\State $TSP = F_{tsp}(TSP,root(TSP))$
\If {data parallel}
\State $TSP = F_{fus}(TSP)$
\EndIf
\State dump(TSP)
\EndProcedure
\end{algorithmic}
\end{algorithm}

%--------------------
\subsection{Implementation choices}

SkelGIS + OpenMP + L2C
